\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-10-03 gwnnDraft02.tex
changes to captions

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The Redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave power spectrum for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity curves for space-based interferometers and pulsar timing arrays as described in Section~\ref{sec:detection}; Lisa (blue), DECIGO (light blue), BBO (red), SKA 5-year (purple), SKA 10-year (orange), SKA 20-year (green). Details of these curves are given in Sec.~\ref{sec:detection}.}
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, more detail in the construction of these curves is given in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sectors contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{highest.png}
\hfill
\includegraphics[width=.45\textwidth]{TwoFar.png} 
\hfill
\includegraphics[width=.45\textwidth]{TwoMed.png} 
\hfill
\includegraphics[width=.45\textwidth]{energydensity.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave power spectrum for various models. The top row consists of two models with only two hidden sectors undergoing SFOPT at differing temperatures. The bottom left (right) are examples of exotic sectors, discussed in section V, with sectors reheated at $T \sim 0.25\,T_{SM}$ ($T \sim 0.17\,T_{SM}$). All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity curves for space-based interferometers and pulsar timing arrays; Lisa~\citep{Audley:2017drz} (blue), DECIO (light blue)~\citep{10.1093/ptep/pty078}, BBO (red)~\citep{PhysRevD.72.083005} , SKA~\cite{Janssen:2014dka} 5-year (purple), SKA 10-year (orange), SKA 20-year (green).   }
\label{fig:Haa}
\end{figure*}

@@

In the former case, this demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors. Individual SFOPT are understood to create GW that are assumed to follow an approximate power law. If a model predicts of the presence of 2, 5, or more additional sectors, or features a single extra sector with multiple PTs, deviations from a standard power law can occur. The multiple transitions that occur in the models outlined here create signals that follow this trend: although the individual GW do obey approximate power laws, their sum does not --- leading to a unique signal indicating so-called dark complexity. Explicitly, a broadening or distortion of the signal around the peak frequency could point to a multi-SFOPT scenario and gently guide us in the direction of multiple hidden sectors, ultimately providing another means to test and constrain theories.

Shifting to the other part of our framework, our results leads to the question ``are current experiments able to probe non power law signals?" For frequency ranges away from the peak of the total GW signals the quoted detection thresholds should hold: the signals fall off in a fashion identical to a power law. However, for areas around the peak frequency the answer is less clear; the PLI curves are built under the assumption of a power law. This points to the need for future work to better understand detection prospects for the next generation of GW detectors.

>>>>

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The Redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave power spectrum for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity curves for space-based interferometers and pulsar timing arrays as described in Section~\ref{sec:detection}; Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red), SKA~\cite{Janssen:2014dka} 5-year (purple), SKA 10-year (orange), SKA 20-year (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, more detail in the construction of these curves is given in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sectors contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

@@

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=.45\textwidth ]{highest.png}
\hfill
\includegraphics[width=.45\textwidth]{TwoFar.png} 
\hfill
\includegraphics[width=.45\textwidth]{TwoMed.png} 
\hfill
\includegraphics[width=.45\textwidth]{energydensity.png} 
\end{minipage}
\hfill
\caption{ Gravitational wave power spectrum for various models. The top row consists of two models with only two hidden sectors undergoing SFOPT at differing temperatures. The bottom left (right) are examples of exotic sectors, discussed in section V, with sectors reheated at $T \sim 0.25\,T_{SM}$ ($T \sim 0.17\,T_{SM}$). All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded curves are the same as Fig.~\ref{fig::Nnatural}.   }
\label{fig:Haa}
\end{figure*}

@@

In the former case, this demonstrates the power of GW signals to probe deep into the unknown arena of complex hidden sectors. Individual SFOPT are understood to create GW that are assumed to follow an approximate power law. If a model predicts of the presence of 2, 5, or more additional sectors, or features a single extra sector with multiple PTs, deviations from a standard power law can occur. The multiple transitions that occur in the models outlined here create signals that follow this trend: although the individual GW do obey approximate power laws, their sum does not --- leading to a unique signal indicating so-called dark complexity. Explicitly, a broadening or distortion of the signal around the peak frequency, precisely where the signal has the most energy, could point to a multi-SFOPT scenario and gently guide us in the direction of multiple hidden sectors, ultimately providing another means to test and constrain theories.

Shifting to the other part of our framework, our results leads to the question ``how well can experiments probe non power law signals?" For frequency ranges away from the peak of the total GW signals the quoted detection thresholds should hold: the signals fall off as a power law to a very good approximation. However, for areas around the peak frequency the answer is less clear; the PLI curves are built under the assumption of a power law. This points to the need for future work to better understand detection prospects for the next generation of GW detectors.
